### 2013-12-01 paper/main.tex
converted ignition points to time in model description
Previously the engine model response with no control
showed rpm versus ignition points.  But this did not
indicate how much time it took to stabilize.
A conversion was found to plot these with time.

\centerline{\Large \textbf{A Survey of Control Systems Applied to}}
\centerline{\Large \textbf{the Idle Control of an Automotive Engine}}
\vspace{0.1in}
\centerline{\normalsize {Jeremiah Mahler}}
\centerline{\small {\href{mailto:jmahler@mail.csuchico.edu}{jmahler@mail.csuchico.edu}} }
\vspace{0.1in}
\centerline{\normalsize {CSU Chico}}
%\centerline{\today}
%\vspace{0.1in}
\centerline{\small \today}
\vspace{0.2in}
\centerline{\LARGE \textbf{DRAFT}}
\vspace{0.2in}

@@

\section{Engine Model}

@@

\begin{figure}[htbp!]
\begin{center}
\includegraphics[scale=0.7]{../octave/engine_model_plot}
\end{center}
\caption{Output response of engine model with no control.
The control can range from zero to one.
Here the control is constant at 0.65 which resulted in a stable
rpm near 900.\protect\footnotemark}
\label{fig:em_plot}
\end{figure}
\footnotetext{The Matlab source to produced these plots is in Appendix \ref{app:em_matlab}.}

\clearpage
\section{Regulator Control System}

@@

% {{{ engine regulator system
\begin{figure}[hbp!]
\begin{center}

@@

% TODO - steady state model

% TODO - open loop performance

@@

\clearpage
\section{Engine Model Matlab Source}
\label{app:em_matlab}

@@

>>>>

\centerline{\Large \textbf{A Survey of Control Systems for Stabilizing}}
\centerline{\Large \textbf{the Idle of a Spark Ignition Engine}}
\vspace{0.1in}
\centerline{\normalsize {Jeremiah Mahler}}
\centerline{\small {\href{mailto:jmahler@mail.csuchico.edu}{jmahler@mail.csuchico.edu}} }
\vspace{0.1in}
\centerline{\normalsize {CSU Chico}}
%\centerline{\today}
%\vspace{0.1in}
\centerline{\small \today}
\vspace{0.2in}
\centerline{\LARGE \textbf{DRAFT}}
\vspace{0.2in}

@@

% {{{ Engine Model
\section{Engine Model}

@@

% {{{ fig:em_plot
\begin{figure}[htbp!]
\begin{center}
\includegraphics[scale=0.7]{../octave/engine_model_plot}
\end{center}
\caption{Output response of engine model with no control.
The control can range from zero to one.
Here the control is constant at 0.65 which resulted in a stable
rpm near 900.\protect\footnotemark}
\label{fig:em_plot}
\end{figure}
\footnotetext{The Matlab source to produced these plots is in Appendix \ref{app:em_matlab}.}
% }}}

It is not apparent from Figure \ref{fig:em_plot} exactly how long
this process takes.
The $x$ axis represents ignition events, not time.
The rpm can be converted to time per ignition event by examining
the units as shown in Equation \ref{eq:rpmtime}.

\begin{align}
\frac{1}{x \frac{\text{rev}}{\text{min}}
			\cdot n \frac{\text{ign}}{\text{rev}}
			\cdot \frac{\text{min}}{60\text{sec}}
		}
&= y \frac{\text{sec}}{\text{ign}} \label{eq:rpmtime} \\
& \notag \\
	\frac{\text{rev}}{\text{min}} &: \text{revolutions per minute (rpm)} \notag \\
	\frac{\text{ign}}{\text{rev}} &: \text{number ignition points per revolution (4 for 8 cylinder engine)} \notag \\
	\frac{\text{sec}}{\text{ign}} &: \text{seconds per ignition point} \notag
\end{align}

Applying this equation results in the response shown
in Figure \ref{fig:rpmtime_plot}.
It can be seen that this plot reveals that it taks over two seconds
for the system to stabilize with no control.
Clearly this is unacceptably slow and it is likely that with
the application of a controller its performance can be improved.

% {{{ fig:rpmtime_plot
\begin{figure}[htbp!]
\begin{center}
\includegraphics[scale=0.7]{../octave/engine_model_rpmtime_plot}
\end{center}
\caption{Engine rpm response compared to time with a constant
0.65 input control.\protect\footnotemark}
\label{fig:rpmtime_plot}
\end{figure}
\footnotetext{The Matlab source of the RPM vs time plot is in Appendix \ref{app:rpmtime_plot_source}.}
% }}}

\clearpage

% {{{ Pole Placement
\section{Pole Placement}

% {{{ Regulator Control System
\clearpage
\section{Regulator Control System}

@@

% {{{ fig:rpm1, Regulator System
\begin{figure}[hbp!]
\begin{center}

@@

@@

\clearpage
\section{Engine Model Matlab Source}
\label{app:em_matlab}
\label{app:rpmtime_plot_source}

@@

\clearpage
\lstinputlisting[
	caption={Matlab source to plot the response of engine model
		versus time.}
	]{../octave/engine_model_rpmtime_plot.m}
